# Import data sources {#data-imports}

## Site list
```{r, read-site-list}
site_list=read.csv(file="site_list.csv")
DT::datatable(site_list, rownames=FALSE, filter="top",
			options = list(scrollY = '600px', paging = FALSE, scrollX=TRUE, dom="ltipr")
)
```

## Lake elevation
```{r, read-elev, cache=T}
elevation=dataRetrieval::readNWISdv(siteNumbers=site_list$SiteNumber[site_list$SiteType=="Elevation gauge"],parameterCd="62614")
elevation=dplyr::rename(elevation, elev_ft=X_62614_00003)
```

## Tributary discharge

### Read discharge gauges
```{r, read-discharge, cache=T}
discharge=dataRetrieval::readNWISuv(siteNumbers=site_list$SiteNumber[site_list$SiteType=="Discharge gauge"],parameterCd="00060", tz="America/Denver", startDate = "2010-01-01", endDate = Sys.Date())
discharge=dplyr::rename(discharge, discharge_cfs=X_00060_00000, )
summary(discharge)
```

### Estimate BRB discharge from Corinne site
```{r, est-brb-dis}
brb_dis=discharge[discharge$site_no==10126000,]
brb_dis$site_no=10010060
brb_dis$discharge_cfs=83.0813+1.0958*brb_dis$discharge_cfs
discharge=rbind(discharge,brb_dis)
```

## Lake & tributary water quality data

```{r, read-wq, cache=T}
wq_data=dataRetrieval::readNWISqw(expanded=F,tz="America/Denver",startDate="01-01-2010", endDate=Sys.Date(),
		siteNumbers=site_list$SiteNumber[site_list$SiteType %in% c("Lake", "Lake, DBL", "Discharge gauge", "Causeway breach")],
		parameterCd=c(
			"00098",			
			"00665",
			"00666",
			"32209",
			"62854",
			"62855",
			"70305",
			"00480",
			"72263",
			"00095")
)

wq_data=dplyr::rename(wq_data,
	Sampling_depth_m=p00098,
	Total_phosphorus_mgL=p00665,
	Filtered_phosphorus_mgL=p00666,
	Chlorophyll_a_ugL=p32209,
	Filtered_nitrogen_mgL=p62854,
	Total_nitrogen_mgL=p62855,
	Salinity_gL=p70305,
	Density_gcm3=p72263,
	SpecCond_uScm=p00095)
```

***Note - removing a few anomolously high data points TP > 8 & TN > 40. These appeared to be outliers based on figures in previous versions.
```{r, rem-anom}
wq_data=within(wq_data, {
	Total_phosphorus_mgL[Total_phosphorus_mgL>8] = NA
	Total_nitrogen_mgL[Total_nitrogen_mgL>40] = NA
})
```

```{r, rel-depth}
## Define surface and bottom samples, add month
wq_data=within(wq_data,{
	rel_depth=NA
	rel_depth[Sampling_depth_m<=1]="Surface"
	rel_depth[Sampling_depth_m>=4]="Bottom"
	month=lubridate::month(sample_dt)
})
```

## Calculate N:P ratios
```{r, np-ratios}
wq_data=within(wq_data,{
	FN_FP=Filtered_nitrogen_mgL/Filtered_phosphorus_mgL * (30.974/14.007)
	TN_TP=Total_nitrogen_mgL/Total_phosphorus_mgL * (30.974/14.007)
})
```


## Read site locations
```{r, read-sites, cache=T}
sites=dataRetrieval::readNWISsite(siteNumbers=site_list$SiteNumber)
```

## Add site metadata to site locations
```{r, site-metadata}
sites=dataRetrieval::readNWISsite(siteNumbers=site_list$SiteNumber)
sites=merge(sites,site_list, by.x=c("site_no","station_nm"), by.y=c("SiteNumber","SiteName"))
```

## Add site metadata to WQ data
```{r, site-metadata-wq}
wq_data=merge(wq_data,sites, all.x=T)
```




