# Gilbert Bay nutrient pool calculations {#gb-nut-pools}
Following the approach from Naftz et al. 2017 to calculate DBL & UBL volumes and constituent masses.

## Calculate UBL & DBL volumes
```{r, volume-calcs}
gb_elev_vol=subset(elevation, site_no==10010000)
gb_elev_vol=within(gb_elev_vol, {
	elev_m=elev_ft/3.28084
	DBL_elev_N=-46.51008+1.03163*elev_m
	DBL_elev_N[DBL_elev_N<1269.43714800001]=1269.43714800001 #Setting minimum possible N DBL elevation at inflection point of formula
	DBL_elev_S=-104.967+1.07643*elev_m
	DBL_elev_S[DBL_elev_S<1270.92582600001]=1270.92582600001 #Setting minimum possible S DBL elevation at inflection point of formula
	DBL_vol_N_m3=59633174862208.5-93949911022.036*DBL_elev_N+37003662.2562751*DBL_elev_N^2
	DBL_vol_S_m3=120180569424473-189400877631.774*DBL_elev_S+74622375.23038100*DBL_elev_S^2
	DBL_vol_total_m3=DBL_vol_S_m3+DBL_vol_N_m3
	UBL_vol_m3=(137054926141519-216113402592.612*elev_m+85193657.0902014*elev_m^2)-DBL_vol_total_m3
	GB_vol_m3=137054926141519-216113402592.612*elev_m+85193657.0902014*elev_m^2
	DBL_vol_acreft=DBL_vol_total_m3*0.000810714
	UBL_vol_acreft=UBL_vol_m3*0.000810714
	GB_vol_acreft=GB_vol_m3*0.000810714
	year=lubridate::year(Date)
	month=lubridate::month(Date)
})

DBLn_vol_YrMo=aggregate(DBL_vol_N_m3~year+month,FUN='mean',gb_elev_vol)
DBLs_vol_YrMo=aggregate(DBL_vol_S_m3~year+month,FUN='mean',gb_elev_vol)
UBL_vol_YrMo=aggregate(UBL_vol_m3~year+month,FUN='mean',gb_elev_vol)
gb_vol_YrMo=merge(UBL_vol_YrMo,DBLn_vol_YrMo,by=c("year","month"))
gb_vol_YrMo=merge(gb_vol_YrMo,DBLs_vol_YrMo,by=c("year","month"))
```

## ID WQ samples as UBL or DBL representatives
Defined DBL samples as those occurring at sites 405356112205601 or 410644112382601 and with sampling depth >= 5 meters (see map).  
Defined UBL samples as any sample in Gilbert Bay >= 2 meters sampling depth. 

```{r, ubl-dbl}
wq_data_flat=within(wq_data_flat, {
	bl=NA
	bl[site_no==405356112205601 & Sampling_depth_m >=5] = 'DBLs'
	bl[site_no==410644112382601 & Sampling_depth_m >=5] = 'DBLn'
	bl[Bay=='Gilbert' & Sampling_depth_m <=2] = 'UBL'
})
```



## Put all constituents in same units (mg/L) to simplify calculations
```{r, pool-units}
gb_pool_wq_data=subset(wq_data_flat, !is.na(bl) & variable %in% c('Salinity_gL','Chlorophyll_a_ugL','Total_phosphorus_mgL','Filtered_phosphorus_mgL','Filtered_nitrogen_mgL','Total_nitrogen_mgL'))
gb_pool_wq_data=droplevels(within(gb_pool_wq_data, {
	variable=as.character(variable)
	value[variable=='Salinity_gL'] = value * 1000
	value[variable=='Chlorophyll_a_ugL'] = value/1000
	variable[variable=='Salinity_gL'] = 'Salinity_mgL'
	variable[variable=='Chlorophyll_a_ugL'] = 'Chlorophyll_a_mgL'
	year=lubridate::year(sample_dt)
	month=lubridate::month(sample_dt)
}))
knitr::kable(table(gb_pool_wq_data$variable))
```

## Calculate monthly means & SDs for UBL & DBL WQ parameters
```{r, calc-mon}
mon_means=aggregate(value~year+month+variable+param+bl, gb_pool_wq_data, FUN='mean')
mon_means=reshape2::dcast(mon_means, year+month+variable+param~bl, value.var='value')
mon_means=dplyr::rename(mon_means, DBL_mean_s=DBLs, DBL_mean_n=DBLn, UBL_mean=UBL)
```

## Merge monthly WQ means & monthly volume estimates
```{r}
gb_pool_data=merge(gb_vol_YrMo, mon_means)
```

## Calculate masses
```{r, calc-mass}
gb_pool_data=within(gb_pool_data, {
	UBL_mass_kg=UBL_vol_m3*UBL_mean*0.001
	DBL_mass_kg=(DBL_mean_n*DBL_vol_N_m3 + DBL_mean_s*DBL_vol_S_m3)*0.001
	tot_mass_kg=DBL_mass_kg+UBL_mass_kg
	YrMo=as.Date(paste0(year,"-",month,"-01"),format='%Y-%m-%d')
})
gb_pool_data=gb_pool_data[order(gb_pool_data$YrMo),]
```

## Compare filtered (dissolved) nitrogen results to Natfz 2017
```{r, jvl-dn-comp1}
jv_dn_mass=subset(gb_pool_data, variable=='Filtered_nitrogen_mgL')
naftz_dn_mass=read.csv(file='naftz2017_DN_mass.csv')
comp=merge(jv_dn_mass, naftz_dn_mass)
plot(tot_mass_kg~DN_total_kg, comp, main='Gilbert Bay dissolved nitrogen pool (kg)', xlab='Natz 2017 estimate', ylab='JVL estimate', cex=1.5)
abline(0,1, lty=2)
plot(DBL_mass_kg~DN_DBL_kg, comp, main='DBL dissolved nitrogen pool (kg)', xlab='Natz 2017 estimate', ylab='JVL estimate', cex=1.5)
abline(0,1, lty=2)
plot(UBL_mass_kg~DN_UBL_kg, comp, main='UBL dissolved nitrogen pool (kg)', xlab='Natz 2017 estimate', ylab='JVL estimate', cex=1.5)
abline(0,1, lty=2)
comp=comp[order(comp$YrMo),]
jvl_dn_comp=plot_ly(data=comp) %>%
	add_trace(type='scatter', y=~tot_mass_kg, x=~YrMo, name='JVL', mode='lines+markers', marker = list(color = cols[[3]][[1]],size = 10,line = list(color = cols[[3]][[2]],width = 2)), line = list(color = cols[[3]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~DN_total_kg, x=~YrMo, name='Naftz', mode='lines+markers', marker = list(symbol=2, color = cols[[6]][[1]],size = 10,line = list(color = cols[[6]][[2]],width = 2)), line = list(color = cols[[6]][[2]],width = 2)) %>%
	layout(title = "Gilbert Bay filtered nitrogen pool (UBL + DBL)",
		xaxis = list(title = ""),
		yaxis = list(side = 'left', title = 'Filtered nitrogen mass (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
jvl_dn_comp

```

## Mass plots for all constituents
### Total nitrogen
```{r, tn-mass-plot}
tn_mass_plot=plot_ly(data=gb_pool_data[gb_pool_data$param=='Total nitrogen',]) %>%
	add_trace(type='scatter', y=~tot_mass_kg, x=~YrMo, mode = 'lines+markers', name='Total', marker = list(color = cols[[3]][[1]],size = 10,line = list(color = cols[[3]][[2]],width = 2)), line = list(color = cols[[3]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~DBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Deep brine layer', name='Deep brine layer', marker = list(symbol=2, color = cols[[6]][[1]],size = 10,line = list(color = cols[[6]][[2]],width = 2)), line = list(color = cols[[6]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~UBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Upper brine layer', marker = list(symbol=4, color = cols[[7]][[1]],size = 10,line = list(color = cols[[7]][[2]],width = 2)),line = list(color = cols[[7]][[2]],width = 2)) %>%
	layout(title = "Gilbert Bay total nitrogen pool",
		xaxis = list(title = "", range=c(min(gb_pool_data$YrMo), max(gb_pool_data$YrMo))),
		yaxis = list(side = 'left', title = 'Total nitrogen mass (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
tn_mass_plot
```

### Filtered (dissolved) nitrogen
```{r, fn-mass-plot}
fn_mass_plot=plot_ly(data=gb_pool_data[gb_pool_data$param=='Filtered nitrogen',]) %>%
	add_trace(type='scatter', y=~tot_mass_kg, x=~YrMo, mode = 'lines+markers', name='Total', marker = list(color = cols[[3]][[1]],size = 10,line = list(color = cols[[3]][[2]],width = 2)), line = list(color = cols[[3]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~DBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Deep brine layer', marker = list(symbol=2, color = cols[[6]][[1]],size = 10,line = list(color = cols[[6]][[2]],width = 2)), line = list(color = cols[[6]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~UBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Upper brine layer', marker = list(symbol=4, color = cols[[7]][[1]],size = 10,line = list(color = cols[[7]][[2]],width = 2)),line = list(color = cols[[7]][[2]],width = 2)) %>%
	layout(title = "Gilbert Bay filtered nitrogen pool",
		xaxis = list(title = "", range=c(min(gb_pool_data$YrMo), max(gb_pool_data$YrMo))),
		yaxis = list(side = 'left', title = 'Filtered nitrogen mass (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
fn_mass_plot
```

### Total phosphorus
```{r, tp-mass-plot}
tp_mass_plot=plot_ly(data=gb_pool_data[gb_pool_data$param=='Total phosphorus',]) %>%
	add_trace(type='scatter', y=~tot_mass_kg, x=~YrMo, mode = 'lines+markers', name='Total', marker = list(color = cols[[3]][[1]],size = 10,line = list(color = cols[[3]][[2]],width = 2)), line = list(color = cols[[3]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~DBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Deep brine layer', marker = list(symbol=2, color = cols[[6]][[1]],size = 10,line = list(color = cols[[6]][[2]],width = 2)), line = list(color = cols[[6]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~UBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Upper brine layer', marker = list(symbol=4, color = cols[[7]][[1]],size = 10,line = list(color = cols[[7]][[2]],width = 2)),line = list(color = cols[[7]][[2]],width = 2)) %>%
	layout(title = "Gilbert Bay total phosphorus pool",
		xaxis = list(title = "", range=c(min(gb_pool_data$YrMo), max(gb_pool_data$YrMo))),
		yaxis = list(side = 'left', title = 'Total phosphorus mass (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
tp_mass_plot
```

### Filtered (dissolved) phosphorus
```{r, fp-mass-plot}
fp_mass_plot=plot_ly(data=gb_pool_data[gb_pool_data$param=='Filtered phosphorus',]) %>%
	add_trace(type='scatter', y=~tot_mass_kg, x=~YrMo, mode = 'lines+markers', name='Total', marker = list(color = cols[[3]][[1]],size = 10,line = list(color = cols[[3]][[2]],width = 2)), line = list(color = cols[[3]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~DBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Deep brine layer', marker = list(symbol=2, color = cols[[6]][[1]],size = 10,line = list(color = cols[[6]][[2]],width = 2)), line = list(color = cols[[6]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~UBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Upper brine layer', marker = list(symbol=4, color = cols[[7]][[1]],size = 10,line = list(color = cols[[7]][[2]],width = 2)),line = list(color = cols[[7]][[2]],width = 2)) %>%
	layout(title = "Gilbert Bay filtered phosphorus pool",
		xaxis = list(title = "", range=c(min(gb_pool_data$YrMo), max(gb_pool_data$YrMo))),
		yaxis = list(side = 'left', title = 'Filtered phosphorus mass (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
fp_mass_plot
```

### Salt
```{r, salt-mass-plot}
salt_mass_plot=plot_ly(data=gb_pool_data[gb_pool_data$param=='Salinity',]) %>%
	add_trace(type='scatter', y=~tot_mass_kg, x=~YrMo, mode = 'lines+markers', name='Total', marker = list(color = cols[[3]][[1]],size = 10,line = list(color = cols[[3]][[2]],width = 2)), line = list(color = cols[[3]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~DBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Deep brine layer', marker = list(symbol=2, color = cols[[6]][[1]],size = 10,line = list(color = cols[[6]][[2]],width = 2)), line = list(color = cols[[6]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~UBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Upper brine layer', marker = list(symbol=4, color = cols[[7]][[1]],size = 10,line = list(color = cols[[7]][[2]],width = 2)),line = list(color = cols[[7]][[2]],width = 2)) %>%
	layout(title = "Gilbert Bay salt pool",
		xaxis = list(title = "", range=c(min(gb_pool_data$YrMo), max(gb_pool_data$YrMo))),
		yaxis = list(side = 'left', title = 'Salt mass (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
salt_mass_plot
```

### Chlorophyll a
```{r, chla-mass-plot}
chla_mass_plot=plot_ly(data=gb_pool_data[gb_pool_data$param=='Chlorophyll a',]) %>%
	add_trace(type='scatter', y=~tot_mass_kg, x=~YrMo, mode = 'lines+markers', name='Total', marker = list(color = cols[[3]][[1]],size = 10,line = list(color = cols[[3]][[2]],width = 2)), line = list(color = cols[[3]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~DBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Deep brine layer', marker = list(symbol=2, color = cols[[6]][[1]],size = 10,line = list(color = cols[[6]][[2]],width = 2)), line = list(color = cols[[6]][[2]],width = 2)) %>%
	add_trace(type='scatter', y=~UBL_mass_kg, x=~YrMo, mode = 'lines+markers', name='Upper brine layer', marker = list(symbol=4, color = cols[[7]][[1]],size = 10,line = list(color = cols[[7]][[2]],width = 2)),line = list(color = cols[[7]][[2]],width = 2)) %>%
	layout(title = "Gilbert Bay chlorophyll a pool",
		xaxis = list(title = "", range=c(min(gb_pool_data$YrMo), max(gb_pool_data$YrMo))),
		yaxis = list(side = 'left', title = 'Chlorophyll a mass (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
chla_mass_plot
```

## Ideas for carrying sampling variability/error through analyses
*Estimate mass by each individual sample, rather than aggregating to monthly means. Use variability 
in these estimates to characterize variability overall





