# Gilbert Bay nutrient load calculations {#gb-nut-loads}
Following the approach from Naftz et al. 2017 to calculate constituent loads to Gilbert Bay.
Note that the BRB outflow discharge was estimated from the Bear River at Corinne gauge as part of the data import process in chapter 2, 'Data imports'.

## Setup
### Flatten & subset to inflow sites
```{r, load-setup1}
load_data_flat=reshape2::melt(wq_data, id.vars=c("site_no", "station_nm", "Bay", "sample_dt", "sample_tm","dec_lat_va","dec_long_va", "month"),
							measure.vars=c("Total_phosphorus_mgL","Filtered_phosphorus_mgL","Filtered_nitrogen_mgL","Total_nitrogen_mgL"),
							na.rm=T)

load_data_flat=subset(load_data_flat, site_no %in% c('411403112200801','410401112134801','10010060','10141000','10172630'))
```

### Merge WQ & discharge data
#### Aggregate to daily values
```{r, load-setup2}
load_data_daily=aggregate(value~site_no+station_nm+sample_dt+variable, data=load_data_flat, FUN='mean')
discharge_daily=aggregate(discharge_cfs~site_no+date, data=discharge, FUN='mean')
names(discharge_daily)[names(discharge_daily)=='date']='sample_dt'
```

#### Merge
```{r, load-setup3}
dim(load_data_daily)
load_data_daily=merge(load_data_daily, discharge_daily)
dim(load_data_daily)
boxplot(discharge_cfs~site_no, load_data_daily)
load_data_daily=subset(load_data_daily, discharge_cfs>0)
```
*** Note - removed records associated with negative discharge values (only occurred at FB outflow site). Unsure how Naftz 2017 dealt with this. This prevents a 'double-counting' of nutrients into Gilbert Bay. ***

#### Loading models
```{r, load-models}
bestMod=function(data){
	return(rloadest::selBestModel(constituent='value', data=data, flow = "discharge_cfs", dates = "sample_dt", conc.units="mg/L", criterion="AIC"))
}
load_models=plyr::dlply(load_data_daily, c('site_no','station_nm','variable'), .fun=bestMod)
```

#### Estimate loads
```{r, load-est}
pred_loads=data.frame(matrix(nrow=0,ncol=9))
for(n in 1:length(load_models)){
	site=strsplit(names(load_models)[n], "[.]")[[1]][1]
	stn=strsplit(names(load_models)[n], "[.]")[[1]][2]
	var=strsplit(names(load_models)[n], "[.]")[[1]][3]
	newdata=subset(load_data_daily, site_no==site, variable=var)
	load=rloadest::predLoad(fit=load_models[[n]], newdata=newdata, by="calendar year", allow.incomplete = T)
	load=within(load, {
		site_no=site
		station_nm=stn
		variable=var
		ann_load_kg=Flux*365.25
	})
	pred_loads=rbind(pred_loads, as.data.frame(load))
}
```

#### Load table
```{r, load-table}
pred_loads[sapply(pred_loads, is.character)] <- lapply(pred_loads[sapply(pred_loads, is.character)], as.factor)
DT::datatable(pred_loads, rownames=FALSE, filter="top",
		options = list(scrollY = '600px', paging = FALSE, scrollX=TRUE, dom="ltipr")) %>%
	DT::formatRound(columns=c(3:8), digits=1)
```

#### Annual DN load
```{r}
ann_dn_load=aggregate(ann_load_kg~Period, subset(pred_loads, variable=='Filtered_nitrogen_mgL'), FUN='sum')
knitr::kable(ann_dn_load)
```

#### Load figures
