# Gilbert Bay nutrient load calculations {#gb-nut-loads}
Following the approach from Naftz et al. 2017 to calculate constituent loads to Gilbert Bay.
Note that the BRB outflow discharge was estimated from the Bear River at Corinne gauge as part of the data import process in chapter 2, 'Data imports'.

## Setup
### Flatten & subset to inflow sites
```{r, load-setup1}
load_data_flat=reshape2::melt(wq_data, id.vars=c("site_no", "station_nm", "Bay", "sample_dt", "sample_tm","dec_lat_va","dec_long_va", "month"),
							measure.vars=c("Total_phosphorus_mgL","Filtered_phosphorus_mgL","Filtered_nitrogen_mgL","Total_nitrogen_mgL"),
							na.rm=T)

load_data_flat=subset(load_data_flat, site_no %in% c('410401112134801','10010060','10141000','10172630'))
```

### Merge WQ & discharge data
#### Aggregate to daily values
```{r, load-setup2}
load_data_daily=aggregate(value~site_no+station_nm+sample_dt+variable, data=load_data_flat, FUN='mean')
discharge_daily=aggregate(discharge_cfs~site_no+date, data=discharge, FUN='mean')
names(discharge_daily)[names(discharge_daily)=='date']='sample_dt'
```

#### Merge
```{r, load-setup3}
dim(load_data_daily)
load_data_daily=merge(load_data_daily, discharge_daily)
dim(load_data_daily)
boxplot(discharge_cfs~site_no, load_data_daily)
load_data_daily=subset(load_data_daily, discharge_cfs>0)
```
*** Note - removed records associated with negative discharge values (only occurred at FB outflow site). Unsure how Naftz 2017 dealt with this. This prevents a 'double-counting' of nutrients into Gilbert Bay. ***

## Loading models
```{r, load-models}
bestMod=function(data){
	return(rloadest::selBestModel(constituent='value', data=data, flow = "discharge_cfs", dates = "sample_dt", conc.units="mg/L", criterion="AIC"))
}
load_models=plyr::dlply(load_data_daily, c('site_no','station_nm','variable'), .fun=bestMod)
```

## Estimate loads
```{r, load-est}
pred_loads=data.frame(matrix(nrow=0,ncol=9))
for(n in 1:length(load_models)){
	site=strsplit(names(load_models)[n], "[.]")[[1]][1]
	stn=strsplit(names(load_models)[n], "[.]")[[1]][2]
	var=strsplit(names(load_models)[n], "[.]")[[1]][3]
	newdata=subset(load_data_daily, site_no==site, variable=var)
	load=rloadest::predLoad(fit=load_models[[n]], newdata=newdata, by="calendar year", allow.incomplete = T)
	load=within(load, {
		site_no=site
		station_nm=stn
		variable=var
		ann_load_kg=Flux*365.25
	})
	pred_loads=rbind(pred_loads, as.data.frame(load))
}
```

## Loading analyses
### Load table
```{r, load-table}
pred_loads[sapply(pred_loads, is.character)] <- lapply(pred_loads[sapply(pred_loads, is.character)], as.factor)
DT::datatable(pred_loads, rownames=FALSE, filter="top",
		options = list(scrollY = '600px', paging = FALSE, scrollX=TRUE, dom="ltipr")) %>%
	DT::formatRound(columns=c(3:8), digits=1)
```

### Annual DN load
```{r, ann-dn-load}
ann_dn_load=aggregate(ann_load_kg~Period, subset(pred_loads, variable=='Filtered_nitrogen_mgL'), FUN='sum')
knitr::kable(ann_dn_load)
```

### Load figures
```{r, load-figures}
total=aggregate(ann_load_kg~variable+Period, pred_loads, FUN='sum')
total$site_no='Total'
pred_loads=plyr::rbind.fill(pred_loads,total)
pred_loads=within(pred_loads, {
	site=NA
	site[site_no=='10010060']='Bear River'
	site[site_no=='10141000']='Weber River'
	site[site_no=='10172630']='Goggin Drain'
	site[site_no=='410401112134801']='Farmington Bay'
	site[site_no=='Total']='Total'
	param=NA
	param[variable=='Total_phosphorus_mgL']='Total phosphorus'
	param[variable=='Filtered_phosphorus_mgL']='Filtered phosphorus'
	param[variable=='Total_nitrogen_mgL']='Total nitrogen'
	param[variable=='Filtered_nitrogen_mgL']='Filtered nitrogen'
	year=as.numeric(gsub('CY ', '', Period))
})

tn_load_plot=
	plot_ly(subset(pred_loads, param=='Total nitrogen'), x=~year) %>%
		add_lines(y=~ann_load_kg, name=~site) %>%
		layout(title = "Gilbert Bay annual total nitrogen loads",
			xaxis = list(title = ""),
			yaxis = list(title = 'Load (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
tn_load_plot


dn_load_plot=
	plot_ly(subset(pred_loads, param=='Filtered nitrogen'), x=~year) %>%
		add_lines(y=~ann_load_kg, name=~site) %>%
		layout(title = "Gilbert Bay annual filtered nitrogen loads",
			xaxis = list(title = ""),
			yaxis = list(title = 'Load (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
dn_load_plot


tp_load_plot=
	plot_ly(subset(pred_loads, param=='Total phosphorus'), x=~year) %>%
		add_lines(y=~ann_load_kg, name=~site) %>%
		layout(title = "Gilbert Bay annual total phosphorus loads",
			xaxis = list(title = ""),
			yaxis = list(title = 'Load (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
tp_load_plot


dp_load_plot=
	plot_ly(subset(pred_loads, param=='Filtered phosphorus'), x=~year) %>%
		add_lines(y=~ann_load_kg, name=~site) %>%
		layout(title = "Gilbert Bay annual filtered phosphorus loads",
			xaxis = list(title = ""),
			yaxis = list(title = 'Load (kg)')
		) %>% 
		config(displaylogo = FALSE,
			modeBarButtonsToRemove = c(
				'sendDataToCloud',
				'hoverClosestCartesian',
				'hoverCompareCartesian',
				'lasso2d',
				'select2d'
			)
		)
dp_load_plot


```


